GEE models

### 2017-10-03 doc/manuscript.Rmd
IS -> insulin sensitivity, analyze -> se

Aim: Our aim was to determine the longitudinal associations of individual non-esterified fatty acids (NEFA) on differences in insulin sensitivity (IS) and beta cell function over 6 years in a cohort of individuals who are at-risk for diabetes.

Methods: In the PROMISE cohort, 477 participants had serum NEFA (mol% and concentration) measured at the baseline visit and completed an OGTT at 3 time points over 6 years. Outcome variables were calculated using the OGTT values. IS was assessed using HOMA2-%S and the Matsuda index, while beta cell function was  assessed using the Insulinogenic index over HOMA-IR (IGI/IR) and the Insulin  Secretion-Sensitivity Index-2 (ISSI-2). Generalized estimating equations were  used, adjusting for time, waist, sex, ethnicity, baseline age, alanine  aminotransferase, and physical activity. NEFA were analyzed as both  concentrations (nmol/mL) and proportions (mol%) of the total fraction.

Results: Participants (73% female, 70% with European-ancestry) had their IS and beta cell function decline by 14-27% over 6 years of follow-up. While in unadjusted models there were several NEFA (e.g. 18:1n-7, 20:1n-9, 22:4n-6) that had negative associations with insulin sensitivity, nearly all of these associations were attenuated in fully adjusted models. In adjusted models, total NEFA, 16:0, 18:1n-9, and 18:2n-6 (as concentrations) had significant negative associations with IGI/IR and ISSI-2, while only 20:5n-3 (as mol%) had positive associations with HOMA2-%S.

>>>>

Aim: Our aim was to determine the longitudinal associations of individual non-esterified fatty acids (NEFA) on differences in insulin sensitivity and beta cell function over 6 years in a cohort of individuals who are at-risk for diabetes.

Methods: In the PROMISE cohort, 477 participants had serum NEFA (mol% and concentration) measured at the baseline visit and completed an OGTT at 3 time points over 6 years. Outcome variables were calculated using the OGTT values. Insulin sensitivity was assessed using HOMA2-%S and the Matsuda index, while beta cell function was  assessed using the Insulinogenic index over HOMA-IR (IGI/IR) and the Insulin  Secretion-Sensitivity Index-2 (ISSI-2). Generalized estimating equations were  used, adjusting for time, waist, sex, ethnicity, baseline age, alanine  aminotransferase, and physical activity. NEFA were analysed as both  concentrations (nmol/mL) and proportions (mol%) of the total fraction.

Results: Participants (73% female, 70% with European-ancestry) had their insulin sensitivity and beta cell function decline by 14-27% over 6 years of follow-up. While in unadjusted models there were several NEFA (e.g. 18:1n-7, 20:1n-9, 22:4n-6) that had negative associations with insulin sensitivity, nearly all of these associations were attenuated in fully adjusted models. In adjusted models, total NEFA, 16:0, 18:1n-9, and 18:2n-6 (as concentrations) had significant negative associations with IGI/IR and ISSI-2, while only 20:5n-3 (as mol%) had positive associations with HOMA2-%S.
